Reinforcement Learning Produces Dominant Strategies for the Iterated Prisoner's Dilemma

### 2017-07-14 main.tex
typos
corrected typos and mainly added `.` at the
figure/table captions.

The Axelrod library now contains many parametrised strategies trained using
machine learning
methods. Most are deterministic, use many rounds of memory, and perform
extremely well in tournaments as will be discussed in Section~\ref{sec:results}.
Training of these strategies will be discussed in Section~\ref{sec:methods}.
These strategies can encode a variety
of other strategies, including classic strategies like Tit For Tat
\cite{Axelrod1980},
handshake strategies, and grudging strategies that always defect after
an opponent defection.

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/lookerup}
    \caption{Diagrammatic representation of the Looker up Archetype}
    \label{fig:lookerup}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/gambler}
    \caption{Diagrammatic representation of the Gambler Archetype}
    \label{fig:gambler}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/ann}
    \caption{Diagrammatic representation of the ANN Archetype}
    \label{fig:ann}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/fsm}
    \caption{Diagrammatic representation of the Finite state machine Archetype}
    \label{fig:fsm}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.5\textwidth}
        \includegraphics[height=.3\textheight]{./assets/FSM16.pdf}
        \caption{Evolved\_FSM\_16: trained to maximise score in a standard
        tournament}
        \label{fig:fsm16}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includegraphics[height=.35\textheight]{./assets/FSM16noise.pdf}
        \caption{Evolved\_FSM\_16\_Noise\_05: trained to maximise score in a
        noisy tournament}
        \label{fig:fsm16noise}
    \end{subfigure}%
    \caption{Trained sixteen state finite state machine players}
    \label{fig:fsm_images}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/hmm}
    \caption{Diagrammatic representation of the Hidden Markov Model Archetype}
    \label{fig:hmm}
\end{figure}

@@

\begin{itemize}
  \item Players are unaware of the number of turns in a match
  \item Players carry no acquired state between matches
  \item Players cannot observe the outcome of other matches
  \item Players cannot identify their opponent by any label or identifier
  \item Players cannot manipulate or inspect their opponents in any way
\end{itemize}

@@

The top 11 performing strategies by median payoff are all strategies trained to maximize
total payoff against a subset of the strategies (Table~\ref{tbl:standard_score}).
The next strategy is Desired Belief Strategy (DBS) \cite{Au2006}.
which actively analyzes the opponent and responds
accordingly. The next two strategies are Winner12, based on a lookup table,
Fool Me Once \cite{axelrodproject}, a grudging strategy that defects indefinitely on
the second defection, and Omega Tit For Tat \cite{kendall2007iterated}.

\begin{table}[!hbtp]
        \centering
        \input{./assets/standard_top_15_scores_summary.tex}
        \caption{Standard Tournament: Mean score per turn of top 15 strategies
            (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).
        The leaderboard is dominated by the trained strategies (indicated by a
        $^{*}$)}
        \label{tbl:standard_score}
\end{table}

@@

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/standard_scores_boxplots.pdf}
        \caption{Standard Tournament: Mean score per turn (ranked by median
        over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments)}
        \label{fig:standard_boxplot}
    \end{figure}
\end{landscape}

@@

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/standard_scores_heatmap.pdf}
    \caption{Standard Tournament: Mean score per turn of row players against
    column players (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments)}
    \label{fig:standard_heatmap}
\end{figure}

The strategies that win the most matches
(Table~\ref{tbl:standard_wins_top_winners} are Defector and Aggravater, followed
by handshaking and zero determinant strategies. This includes two handshaking
strategies that were the result of training to maximize Moran process fixation
(TF1 and TF2). No strategies were trained specifically to win matches. None of
the top scoring strategies appear in the top 15 list of strategies ranked by
match wins. This can be seen in Figure~\ref{fig:standard_winplot} where the
distribution of the number of wins of each strategy is shown.

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_winners_wins_summary.tex}
        \caption{Standard Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median wins over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments)}
        \label{tbl:standard_wins_top_winners}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/standard_wins_boxplots.pdf}
        \caption{Standard Tournament: number of wins per tournament (ranked by
        median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments)}
        \label{fig:standard_winplot}
    \end{figure}
\end{landscape}

The number of wins of the top strategies of Table~\ref{tbl:standard_wins} are
shown in Table~\ref{tbl:standard_wins}. It is evident that although these
strategies score highly they do not win many matches: the strategy with the most
number of wins is the Evolved FSM 16 strategy that at most won 60
(\(60/175\approx34\%\)) matches in a given tournament.

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_wins_summary.tex}
        \caption{Standard Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median score over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments)}
        \label{tbl:standard_wins}
\end{table}

@@

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_ranks_summary.tex}
        \caption{Standard Tournament: Rank in each tournament
        of top 15 strategies (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments)}
        \label{tbl:standard_ranks}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/standard_ranks_boxplots.pdf}
        \caption{Standard Tournament: rank in each tournament (ranked by
        median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments)}
        \label{fig:standard_ranks_boxplot}
    \end{figure}
\end{landscape}

@@

    \caption{Comparison of cooperation rates for Standard Tournament Top 3
    (over 10000 repetitions)}
    \label{fig:comparison_cooperation_heatmaps_standard}
\end{figure}

@@

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/noisy_scores_boxplots.pdf}
        \caption{Noisy (5\%) Tournament: Mean score per turn (ranked by median
        over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments)}
        \label{fig:noisy_score}
    \end{figure}
\end{landscape}

@@

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/noisy_scores_heatmap.pdf}
    \caption{Noisy (5\%) Tournament: Mean score per turn of row players against
    column players (ranked by median over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments)}
    \label{fig:noisy_heatmap}
\end{figure}

As shown in Table~\ref{tbl:noisy_wins_top_winners} and
Figure~\ref{fig:noisy_winplot} the strategies tallying the most wins are
somewhat similar to the standard tournaments, with Defector, the handshaking
CollectiveStrategy, and Aggravate appearing as the top three again.

\begin{table}[!hbtp]
    \centering
        \input{./assets/noisy_top_15_winners_wins_summary.tex}
        \caption{Noisy (5\%) Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median wins over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments)}
        \label{tbl:noisy_wins_top_winners}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/noisy_wins_boxplots.pdf}
        \caption{Noisy (5\%) Tournament: number of wins per tournament (ranked by
        median over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments)}
        \label{fig:noisy_winplot}
    \end{figure}
\end{landscape}

As shown in Table~\ref{tbl:noisy_wins}, the top ranking strategies win a larger
number of matches in the presence of noise. For example Spiteful Tit For Tat in
one tournament won almost all its matches (167).

\begin{table}[!hbtp]
    \centering
        \input{./assets/noisy_top_15_wins_summary.tex}
        \caption{Noisy (5\%) Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median score over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments)}
        \label{tbl:noisy_wins}
\end{table}

@@

\begin{table}[!hbtp]
    \centering
        \input{./assets/noisy_top_15_ranks_summary.tex}
        \caption{Noisy (5\%) Tournament: Rank in each tournament
        of top 15 strategies (ranked by median over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments)}
        \label{tbl:noisy_ranks}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/noisy_ranks_boxplots.pdf}
        \caption{Noisy (5\%) Tournament: rank in each tournament (ranked by
        median over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments)}
        \label{fig:noisy_ranks_boxplot}
    \end{figure}
\end{landscape}

@@

    \caption{Comparison of cooperation rates for the Noisy (5\%) Tournament Top 3
    (over 10000 standard tournaments)}
    \label{fig:comparison_cooperation_heatmaps_noisy}
\end{figure}

>>>>

The Axelrod library now contains many parametrised strategies trained using
machine learning
methods. Most are deterministic, use many rounds of memory, and perform
extremely well in tournaments as will be discussed in Section~\ref{sec:results}.
Training of these strategies will be discussed in Section~\ref{sec:methods}.
These strategies can encode a variety
of other strategies, including classic strategies like Tit For Tat
\cite{Axelrod1980},
handshake strategies, and grudging strategies, that always defect after
an opponent defection.

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/lookerup}
    \caption{Diagrammatic representation of the Looker up Archetype.}
    \label{fig:lookerup}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/gambler}
    \caption{Diagrammatic representation of the Gambler Archetype.}
    \label{fig:gambler}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/ann}
    \caption{Diagrammatic representation of the ANN Archetype.}
    \label{fig:ann}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/fsm}
    \caption{Diagrammatic representation of the Finite state machine Archetype.}
    \label{fig:fsm}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.5\textwidth}
        \includegraphics[height=.3\textheight]{./assets/FSM16.pdf}
        \caption{Evolved\_FSM\_16: trained to maximise score in a standard
        tournament}
        \label{fig:fsm16}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includegraphics[height=.35\textheight]{./assets/FSM16noise.pdf}
        \caption{Evolved\_FSM\_16\_Noise\_05: trained to maximise score in a
        noisy tournament}
        \label{fig:fsm16noise}
    \end{subfigure}%
    \caption{Trained sixteen state finite state machine players.}
    \label{fig:fsm_images}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/hmm}
    \caption{Diagrammatic representation of the Hidden Markov Model Archetype.}
    \label{fig:hmm}
\end{figure}

@@

\begin{itemize}
  \item Players are unaware of the number of turns in a match.
  \item Players carry no acquired state between matches.
  \item Players cannot observe the outcome of other matches.
  \item Players cannot identify their opponent by any label or identifier.
  \item Players cannot manipulate or inspect their opponents in any way.
\end{itemize}

@@

The top 11 performing strategies by median payoff are all strategies trained to maximize
total payoff against a subset of the strategies (Table~\ref{tbl:standard_score}).
The next strategy is Desired Belief Strategy (DBS) \cite{Au2006},
which actively analyzes the opponent and responds
accordingly. The next two strategies are Winner12, based on a lookup table,
Fool Me Once \cite{axelrodproject}, a grudging strategy that defects indefinitely on
the second defection, and Omega Tit For Tat \cite{kendall2007iterated}.

\begin{table}[!hbtp]
        \centering
        \input{./assets/standard_top_15_scores_summary.tex}
        \caption{Standard Tournament: Mean score per turn of top 15 strategies
            (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).
        The leaderboard is dominated by the trained strategies (indicated by a
        $^{*}$).}
        \label{tbl:standard_score}
\end{table}

@@

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/standard_scores_boxplots.pdf}
        \caption{Standard Tournament: Mean score per turn (ranked by median
        over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{fig:standard_boxplot}
    \end{figure}
\end{landscape}

@@

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/standard_scores_heatmap.pdf}
    \caption{Standard Tournament: Mean score per turn of row players against
    column players (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
    \label{fig:standard_heatmap}
\end{figure}

The strategies that win the most matches
(Table~\ref{tbl:standard_wins_top_winners}) are Defector~\cite{Axelrod1984} and Aggravater~\cite{axelrodproject}, followed
by handshaking and zero determinant strategies~\cite{Press2012}. 
This includes two handshaking
strategies that were the result of training to maximize Moran process fixation
(TF1 and TF2). No strategies were trained specifically to win matches. None of
the top scoring strategies appear in the top 15 list of strategies ranked by
match wins. This can be seen in Figure~\ref{fig:standard_winplot} where the
distribution of the number of wins of each strategy is shown.

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_winners_wins_summary.tex}
        \caption{Standard Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median wins over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{tbl:standard_wins_top_winners}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/standard_wins_boxplots.pdf}
        \caption{Standard Tournament: number of wins per tournament (ranked by
        median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{fig:standard_winplot}
    \end{figure}
\end{landscape}

The number of wins of the top strategies of Table~\ref{tbl:standard_wins_top_winners} are
shown in Table~\ref{tbl:standard_wins}. It is evident that although these
strategies score highly they do not win many matches: the strategy with the most
number of wins is the Evolved FSM 16 strategy that at most won 60
(\(60/175\approx34\%\)) matches in a given tournament.

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_wins_summary.tex}
        \caption{Standard Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median score over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{tbl:standard_wins}
\end{table}

@@

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_ranks_summary.tex}
        \caption{Standard Tournament: Rank in each tournament
        of top 15 strategies (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{tbl:standard_ranks}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/standard_ranks_boxplots.pdf}
        \caption{Standard Tournament: rank in each tournament (ranked by
        median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{fig:standard_ranks_boxplot}
    \end{figure}
\end{landscape}

@@

    \caption{Comparison of cooperation rates for Standard Tournament Top 3
    (over 10000 repetitions).}
    \label{fig:comparison_cooperation_heatmaps_standard}
\end{figure}

@@

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/noisy_scores_boxplots.pdf}
        \caption{Noisy (5\%) Tournament: Mean score per turn (ranked by median
        over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments).}
        \label{fig:noisy_score}
    \end{figure}
\end{landscape}

@@

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/noisy_scores_heatmap.pdf}
    \caption{Noisy (5\%) Tournament: Mean score per turn of row players against
    column players (ranked by median over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments).}
    \label{fig:noisy_heatmap}
\end{figure}

As shown in Table~\ref{tbl:noisy_wins_top_winners} and
Figure~\ref{fig:noisy_winplot} the strategies tallying the most wins are
somewhat similar to the standard tournaments, with Defector, the handshaking
CollectiveStrategy~\cite{Li2009}, and Aggravate appearing as the top three again.

\begin{table}[!hbtp]
    \centering
        \input{./assets/noisy_top_15_winners_wins_summary.tex}
        \caption{Noisy (5\%) Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median wins over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments).}
        \label{tbl:noisy_wins_top_winners}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/noisy_wins_boxplots.pdf}
        \caption{Noisy (5\%) Tournament: number of wins per tournament (ranked by
        median over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments).}
        \label{fig:noisy_winplot}
    \end{figure}
\end{landscape}

As shown in Table~\ref{tbl:noisy_wins}, the top ranking strategies win a larger
number of matches in the presence of noise. For example Spiteful Tit For Tat~\cite{Prison1998} 
in one tournament won almost all its matches (167).

\begin{table}[!hbtp]
    \centering
        \input{./assets/noisy_top_15_wins_summary.tex}
        \caption{Noisy (5\%) Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median score over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments).}
        \label{tbl:noisy_wins}
\end{table}

@@

\begin{table}[!hbtp]
    \centering
        \input{./assets/noisy_top_15_ranks_summary.tex}
        \caption{Noisy (5\%) Tournament: Rank in each tournament
        of top 15 strategies (ranked by median over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments).}
        \label{tbl:noisy_ranks}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/noisy_ranks_boxplots.pdf}
        \caption{Noisy (5\%) Tournament: rank in each tournament (ranked by
        median over
        \protect\input{./assets/noisy_number_of_repetitions.tex}tournaments).}
        \label{fig:noisy_ranks_boxplot}
    \end{figure}
\end{landscape}

@@

    \caption{Comparison of cooperation rates for the Noisy (5\%) Tournament Top 3
    (over 10000 standard tournaments).}
    \label{fig:comparison_cooperation_heatmaps_noisy}
\end{figure}
